---
title: "CRDSIAnalysis"
author: "Natalie Levesque"
date: "5/10/2020"
output: html_document
---
I will be analyzing this data for my masters thesis. Due to the small amount of data I was actually able to run on the IRMS, --I will only be looking at the plant community.-- my data is not ideal.

#Setup
##Library
```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(simmr)
library(viridis)
library(psych)
library(lubridate)

```

##Setting up the final dataset
```{r}
#reading in the actual data file I'll be using
producers<-read.csv("ProducersSI.csv")
producers
str(producers)
producers$date<-as.Date.character(producers$date, "%m/%d/%y")
unique(producers$sampletype) #need to go in manually and change some of these later
unique(producers$pond)

#Need to remove Strawberry and Eyak South from the dataset as they aren't replicated.
producers<-subset(producers, pond != "Strawberry")
producers<-subset(producers, pond != "Eyak South")

#Many of the NAs are caused by detritus, so I went in manually and made detritus the name at each taxonomic level.
```

###Pulling a new column "year" out of date
```{r}
producers$year<-producers$date
producers$year<-year(producers$year)
```

#Basic Data Exploration
```{r}
names(producers)
Ccorr.plot<-producers[,c(15,2,3,12,14)]
Ncorr.plot<-producers[,c(16,2,3,12,14)]
pairs.panels(Ccorr.plot, ellipses = F)
pairs.panels(Ncorr.plot, ellipses = F)

```

#Pond Biplots
##Ponds Table
```{r}
pond.summary.table <- producers %>% group_by(pond) %>% 
  summarise(count = length(pond),
  meandC13 = mean(dC13), SdC13 =  sd(dC13), CR = (max(dC13)-min(dC13)),
  meandN15 = mean(dN15), SdN15 =  sd(dN15), NR = (max(dN15)-min(dN15)))

kable(pond.summary.table, digits = 4)

```
##Range Bar Charts
```{r}
CarbonR<-ggplot(data = pond.summary.table, aes(reorder(pond, -CR), CR)) +
  geom_bar(stat="identity")+
  labs(x="Site", 
       y=expression(paste(delta^{13}, " C Range (\u2030)")))+
  theme_classic()+
  theme(axis.text.x = element_text(size = 11, angle = 18, hjust = 0.75))
  
CarbonR

NitrogenR<-ggplot(data = pond.summary.table, aes(reorder(pond, -NR), NR))+
  geom_bar(stat = "identity")+
  labs(x="Site", 
       y=expression(paste(delta^{15}, " N Range (\u2030)")))+
  theme_classic()+
  theme(axis.text.x = element_text(size = 11, angle = 18, hjust = 0.75))
  
NitrogenR
```

##Ponds Graph
```{r, include=FALSE}


# Simple Plot
first.plot <- ggplot(data = producers, aes(dC13, dN15)) +
  geom_point(aes(color = pond), size = 3)+
  ylab(expression(paste(delta^{15}, "N (\u2030)")))+
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=15))

# Ellipse Plot
#assign ellipse size
p.ell <- 0.50

ellipse.plot <- first.plot + 
  stat_ellipse(aes(group = pond, 
                   fill = pond, 
                   color = pond), 
               alpha = 0.3, 
               level = p.ell,
               type = "norm",
               geom = "polygon")

# Error bar plot
errorbar.plot<-first.plot + 
  geom_point(data = pond.summary.table,
             aes(meandC13, meandN15,fill = pond), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7) +
  geom_errorbar(data = pond.summary.table, 
                mapping = aes(x = meandC13, y = meandN15,
                              ymin = meandN15 - 1.96*SdN15, 
                              ymax = meandN15 + 1.96*SdN15), 
                width = 0) +
  geom_errorbarh(data = pond.summary.table, 
                 mapping = aes(x = meandC13, y = meandN15,
                               xmin = meandC13 - 1.96*SdC13,
                               xmax = meandC13 + 1.96*SdC13),
                 height = 0)
```

Plots:

```{r}
first.plot
ellipse.plot
```

#Year Biplots
##Year Table
```{r}
year.summary.table <- producers %>% group_by(year) %>% 
  summarise(count = length(year),
  meandC13 = mean(dC13), SdC13 =  sd(dC13), CR = (max(dC13)-min(dC13)), 
  meandN15 = mean(dN15), SdN15 =  sd(dN15), NR = (max(dN15)-min(dN15)),)

kable(year.summary.table, digits = 4)

```
##Range Bar Charts: Year
```{r}
CarbonR<-ggplot(data = year.summary.table, aes(reorder(year, year), CR)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=count, vjust = -0.25))+
  labs(x="Year", 
       y=expression(paste(delta^{13}, " C Range (\u2030)")))+
  theme_classic()+
  theme(axis.text.x = element_text(size = 11))
  
CarbonR

NitrogenR<-ggplot(data = year.summary.table, aes(reorder(year, year), NR))+
  geom_bar(stat = "identity")+
  geom_text(aes(label=count, vjust = -0.25))+
  labs(x="Year", 
       y=expression(paste(delta^{15}, " N Range (\u2030)")))+
  theme_classic()+
  theme(axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 11))
  
NitrogenR
```
##Years Graph
```{r, include=FALSE}
# Simple Trophic Plot
first.plot <- ggplot(data = producers, aes(dC13, dN15)) +
  geom_point(aes(color = year), size = 3)+
  ylab(expression(paste(delta^{15}, "N (\u2030)")))+
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=15))

# Ellipse Plot for Trophic Groups
#assign ellipse size
p.ell <- 0.50

ellipse.plot <- first.plot + 
  stat_ellipse(aes(group = year, 
                   fill = year, 
                   color = year), 
               alpha = 0.3, 
               level = p.ell,
               type = "norm",
               geom = "polygon")

# Error bar plot
errorbar.plot<-first.plot + 
  geom_point(data = year.summary.table,
             aes(meandC13, meandN15,fill = year), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7) +
  geom_errorbar(data = year.summary.table, 
                mapping = aes(x = meandC13, y = meandN15,
                              ymin = meandN15 - 1.96*SdN15, 
                              ymax = meandN15 + 1.96*SdN15), 
                width = 0) +
  geom_errorbarh(data = year.summary.table, 
                 mapping = aes(x = meandC13, y = meandN15,
                               xmin = meandC13 - 1.96*SdC13,
                               xmax = meandC13 + 1.96*SdC13),
                 height = 0)
```

Plots:

```{r}
first.plot
ellipse.plot

```



#Genus Biplots
##Genus Table
```{r}
genus.summary.table <- producers %>% group_by(genus) %>% 
  summarise(count = length(genus),
  meandC13 = mean(dC13), SdC13 =  sd(dC13), CR = (max(dC13)-min(dC13)),
  meandN15 = mean(dN15), SdN15 =  sd(dN15), NR = (max(dN15)-min(dN15)),)

kable(genus.summary.table, digits = 4)

```
##Range Bar Charts: Genus
```{r}
CarbonR<-ggplot(data = genus.summary.table, aes(reorder(genus, -CR), CR)) +
  geom_bar(stat="identity")+
  labs(x="Genus", 
       y=expression(paste(delta^{13}, " C Range (\u2030)")))+
  theme_classic()+
  theme(axis.text.x = element_text(size = 11, angle = 18, hjust = 0.75))
  
CarbonR

NitrogenR<-ggplot(data = genus.summary.table, aes(reorder(genus, -NR), NR))+
  geom_bar(stat = "identity")+
  labs(x="Genus", 
       y=expression(paste(delta^{15}, " N Range (\u2030)")))+
  theme_classic()+
  theme(axis.text.x = element_text(size = 11, angle = 18, hjust = 0.75))
  
NitrogenR
```
##Genus Graph
```{r, include=FALSE}
# Simple Plot
first.plot <- ggplot(data = producers, aes(dC13, dN15)) +
  geom_point(aes(color = genus), size = 3)+
  ylab(expression(paste(delta^{15}, "N (\u2030)")))+
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=15))

# Ellipse Plot
#assign ellipse size
p.ell <- 0.50

ellipse.plot <- first.plot + 
  stat_ellipse(aes(group = genus, 
                   fill = genus, 
                   color = genus), 
               alpha = 0.3, 
               level = p.ell,
               type = "norm",
               geom = "polygon")

# Error bar plot
errorbar.plot<-first.plot + 
  geom_point(data = genus.summary.table,
             aes(meandC13, meandN15,fill = genus), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7) +
  geom_errorbar(data = genus.summary.table, 
                mapping = aes(x = meandC13, y = meandN15,
                              ymin = meandN15 - 1.96*SdN15, 
                              ymax = meandN15 + 1.96*SdN15), 
                width = 0) +
  geom_errorbarh(data = genus.summary.table, 
                 mapping = aes(x = meandC13, y = meandN15,
                               xmin = meandC13 - 1.96*SdC13,
                               xmax = meandC13 + 1.96*SdC13),
                 height = 0)
```

Plots:

```{r message=FALSE, warning=FALSE}
first.plot
ellipse.plot
```

#Family Biplots
####Genus had too many NAs (56!) So i'm going up a level to increase sample size.
##Family Table
```{r}
family.summary.table <- producers %>% group_by(family) %>% 
  summarise(count = length(family),
  meandC13 = mean(dC13), SdC13 =  sd(dC13), CR = (max(dC13)-min(dC13)),
  meandN15 = mean(dN15), SdN15 =  sd(dN15), NR = (max(dN15)-min(dN15)),)

kable(family.summary.table, digits = 4)

```
##Range Bar Charts: Family
```{r}
#Added the sample size above each barplot()

CarbonR<-ggplot(data = family.summary.table, aes(reorder(family, -CR), CR)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=count, vjust = -0.25))+
  labs(x="Family", 
       y=expression(paste(delta^{13}, " C Range (\u2030)")))+
  theme_classic()+
  theme(axis.text.x = element_text(size = 11, angle = 30, hjust = 0.85))
  
CarbonR

NitrogenR<-ggplot(data = family.summary.table, aes(reorder(family, -NR), NR))+
  geom_bar(stat = "identity")+
  geom_text(aes(label=count, vjust = -0.25))+
  labs(x="Family", 
       y=expression(paste(delta^{15}, " N Range (\u2030)")))+
  theme_classic()+
  theme(axis.text.x = element_text(size = 11, angle = 30, hjust = 0.85))
  
NitrogenR
```

##Family Graph
```{r, include=FALSE}
# Simple Plot
first.plot <- ggplot(data = producers, aes(dC13, dN15)) +
  geom_point(aes(color = family), size = 3)+
  ylab(expression(paste(delta^{15}, "N (\u2030)")))+
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=15))

# Ellipse Plot
#assign ellipse size
p.ell <- 0.50

ellipse.plot <- first.plot + 
  stat_ellipse(aes(group = family, 
                   fill = family, 
                   color = family), 
               alpha = 0.10, 
               level = p.ell,
               type = "norm",
               geom = "polygon")

# Error bar plot
errorbar.plot<-first.plot + 
  geom_point(data = family.summary.table,
             aes(meandC13, meandN15,fill = family), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7) +
  geom_errorbar(data = family.summary.table, 
                mapping = aes(x = meandC13, y = meandN15,
                              ymin = meandN15 - 1.96*SdN15, 
                              ymax = meandN15 + 1.96*SdN15), 
                width = 0) +
  geom_errorbarh(data = family.summary.table, 
                 mapping = aes(x = meandC13, y = meandN15,
                               xmin = meandC13 - 1.96*SdC13,
                               xmax = meandC13 + 1.96*SdC13),
                 height = 0)
```

Family Plots:

```{r message=FALSE, warning=FALSE}
first.plot
ellipse.plot
```

#Sample Type Biplots
##Sample Type Table
```{r}
sampletype.summary.table <- producers %>% group_by(sampletype) %>% 
  summarise(count = length(sampletype),
  meandC13 = mean(dC13), SdC13 =  sd(dC13), 
  meandN15 = mean(dN15), SdN15 =  sd(dN15))

```

##Sample Type Graph
```{r warning=FALSE}
# Simple Plot
first.plot <- ggplot(data = producers, aes(dC13, dN15)) +
  geom_point(aes(color = sampletype), size = 3)+
  ylab(expression(paste(delta^{15}, "N (\u2030)")))+
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=15))

# Ellipse Plot
#assign ellipse size
p.ell <- 0.50

ellipse.plot <- first.plot + 
  stat_ellipse(aes(group = sampletype, 
                   fill = sampletype, 
                   color = sampletype), 
               alpha = 0.3, 
               level = p.ell,
               type = "norm",
               geom = "polygon")

ellipse.plot
```

